---
title: "plots"
format: html
---

```{r}
library(tidyverse)
butter <- read_csv(here::here("butter/data/recipes_butter.csv"))
```

Plot Ideas:
- what recipe has the most butter, which one has the least
- is more butter (per serving) better?
- how many recipes had butter
- are butter forward recipe published more in the winter season?
- has the number of recipes with butter increased or decreased over time (look at recipes with and without)
- has the amount of butter used in recipes decreased over time (in just recipes with butter)
- which author is the most butter forward
- distribution of recipes ratings with and without butter, two plots
- something with the total number of other ingredients, do recipes without butter have more or less total ingredients than those without
- there will probably be something with dessert vs meals IDK what yet though


- Color coded for above or below average for amount of butter
- maybe like the stacked chart of circles I saved in my pins and bin the ratings and color coordinate the amount of butter, below can be a gradient of one color and above and below another color


Basic
```{r}
hist(butter$tablespoons)

just_butter <- butter %>% 
  filter(tablespoons > 0)

ggplot(just_butter) +
  geom_point(aes(x=avg_rating, y = butter_per_serving))

ggplot(just_butter) +
  geom_bar(aes(x=avg_rating, y = butter_per_serving), stat = "identity")
```





```{r}


just_butter_counts <- left_join(just_butter, buttered_author, by = "author")

max_recipes <- max(just_butter_counts$total_recipes, na.rm = TRUE)
max_butter <- max(just_butter_counts$butter_per_serving, na.rm = TRUE)
scaler <- max_recipes / max_butter

ggplot(just_butter_counts, aes(y = author)) +
  # 1. The Bars (placed first so they are "behind")
  # We multiply the count by the scaler to fit the butter axis
  geom_col(aes(x = total_recipes * scaler), 
           fill = "gray90", color = NA) + 
  geom_point(aes(x = butter_per_serving), 
             alpha = 0.6, color = "black") +
  scale_x_continuous(
    name = "Butter per Serving",
    sec.axis = sec_axis(~ . / scaler, name = "Total Number of Recipes")
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6)) # Shrink text if you have many authors

```
















## Does an increase in butter per serving increase the rating (stats)
```{r}
lm_per_serv <- lm(avg_rating~butter_per_serving, data=just_butter)
summary(lm_per_serv)

lm_per_recipe <- lm(avg_rating~tablespoons, data=just_butter)
summary(lm_per_recipe)


spearman_test <- cor.test(just_butter$butter_per_serving, just_butter$avg_rating,
                          method = "spearman")
print(spearman_test) ## p-value is significant, the butter per serving does significantly impact the average rating of the recipe

rho_val <- spearman_test$estimate

ggplot(just_butter, aes(x = butter_per_serving, y = avg_rating)) +
  geom_point(alpha = 0.5) +  # Alpha helps if points overlap
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Relationship between Butter and Rating",
       subtitle = paste("Spearman's ρ =", round(rho_val, 3)),
       x = "Butter per Serving",
       y = "Average Rating")
```